---
title: "IBM HR Analytics"
author: "Himanshu Singh"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly I font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```

# IBM HR Analytics Dataset: An Intro

I will analyse a data set on Human Resoruce Analytics. The [IBM HR Analytics Employee Attrition & Performance data set](https://www.kaggle.com/pavansubhasht/ibm-hr-analytics-attrition-dataset) is a fictional data set created by IBM data scientists.  Among other things, the data set includes employees' income, their distance from work, their position in the company, their level of education, etc. A full description can be found on the website.



```{r , echo = FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(ggthemes)
library(kableExtra)
library(ggrepel)
```

First let us load the dataset.

```{r}

hr_dataset <- read_csv("data/datasets_1067_1925_WA_Fn-UseC_-HR-Employee-Attrition.csv")
glimpse(hr_dataset)

```

I am going to clean the data set, as variable names are in capital letters, some variables are not really necessary, and some variables, e.g., `education` are given as a number rather than a more useful description

`
```{r eval =  FALSE }
hr_cleaned <- hr_dataset %>% 
  clean_names() %>% 
  mutate(
    education = case_when(
      education == 1 ~ "Below College",
      education == 2 ~ "College",
      education == 3 ~ "Bachelor",
      education == 4 ~ "Master",
      education == 5 ~ "Doctor"
    ),
    environment_satisfaction = case_when(
      environment_satisfaction == 1 ~ "Low",
      environment_satisfaction == 2 ~ "Medium",
      environment_satisfaction == 3 ~ "High",
      environment_satisfaction == 4 ~ "Very High"
    ),
    job_satisfaction = case_when(
      job_satisfaction == 1 ~ "Low",
      job_satisfaction == 2 ~ "Medium",
      job_satisfaction == 3 ~ "High",
      job_satisfaction == 4 ~ "Very High"
    ),
    performance_rating = case_when(
      performance_rating == 1 ~ "Low",
      performance_rating == 2 ~ "Good",
      performance_rating == 3 ~ "Excellent",
      performance_rating == 4 ~ "Outstanding"
    ),
    work_life_balance = case_when(
      work_life_balance == 1 ~ "Bad",
      work_life_balance == 2 ~ "Good",
      work_life_balance == 3 ~ "Better",
      work_life_balance == 4 ~ "Best"
    )
  ) %>% 
  select(age, attrition, daily_rate, department,
         distance_from_home, education,
         gender, job_role,environment_satisfaction,
         job_satisfaction, marital_status,
         monthly_income, num_companies_worked, percent_salary_hike,
         performance_rating, total_working_years,
         work_life_balance, years_at_company,
         years_since_last_promotion)


```


After the data cleaning, there are 10 variables whose type are character, including “`attrition`”, “`department`”, “`education`”,“`gender`”, “`job_role`”, “`environment_satisfaction`”, “`job_satisfaction`”, “`marital_status`”, “`performance_rating`”* and “`work_life_balance`”. There are also 9 variables whose type is integer, including “`age`”, “`daily_rate`”, “`distance_from_home`”, “`monthly_income`”, “`nun_companies_worked`”, “`percent_salary_hike`”, “`total_working_years`”, “`years_at_company`” and “`years_since_last_promotion`”. Luckily no missing values exist. 


# Attrition Behaviour
The attrition rate of the copmany is given by the number of employees who have left the company divided by the total number of employees. As shown below, the attrition rate calculated from this dataset is **16.1%.**. 











